Influence of galaxy rotation and outflows in the Lyman Alpha spectral line

### 2017-02-17 paper/paper.tex
updates

Galaxies are ellipsoidal. However, for simplicity we model them as a sphere. This shape facilitates the interpretation of the results in our numerical experiments. Furthermore, this approximation is commonly used in the literature, as it explains a wide variety of observational features (\cite{Ahn03}, \cite{Verhamme06}, \cite{Dijkstra06}). \\

@@

In this model, the LAE has a bulk velocity which is the superposition of the rotation and the outflow, as shown in Fig. \ref{fig:model}. The velocity components are written in Eqs. (\ref{eq:vx}) (\ref{eq:vy}) (\ref{eq:vz}). In these equations, $R$ is the radius of the sphere; $x$, $y$ and $z$ are the coordinates in a cartesian frame; and the $\mp$ signs in $v_x$ and $v_y$ indicate the direction of rotation, respectively. This rotation is a solid body rotation and its direction goes according to the right hand rule applied to the $\hat{k}$ unit vector. The outflow velocity is dependent on the position relative to the center of the galaxy, being it zero at the center and maximum at the edge of the sphere.\\

@@

If $x < 0$, the final frequency $\nu < \nu_{\rm \alpha}$. This translates to the final wavelength being larger than the \lya natural one. The phenomenon is called a redshift in frequency. If, on the contrary, $x > 0$, the photon suffers a blueshift in frequency. \\

@@

The initial emission of photons is taken at the center of the sphere for practicality due to the fact that both, center and off-center emissions, give analogous results. From here, $100000$ photons are emitted with the natural \lya wavelength. Then, they start to behave as described in sub-section \ref{subsec:photons_path_clara} (See Fig. \ref{fig:radiative_transfer}). When each photon is re-emitted, its new wavelength depends on the H atom's velocity (both thermal and bulk) and direction (both initial and final). However the photon's new direction of propagation is random in the rest frame of the atom. \\

@@

The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation. Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires. Depending on \tauh, we need a different number of processors in order to minimize the time demands. For \tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively. The running times go from 7 hours to 1 week approximately, depending on the case. \\

@@

\subsection{Parameters' Initial Values}

In order to define the ranges of \tauh, \vrot and \vout, it is necessary to refer to observational constraints. The common values for typical LAEs are in Tab. \ref{tab:values}. We run CLARA for all the permutations of these 3 parameters. \\

@@

The resulting sets of spectra are in Figs. \ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$ available in the Appendix \ref{ap:results_figures}. We do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms. For this reason, we limit in all the plots \vout to a maximum value of 25 \kms. Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{ap:additional_figures} in Figs. \ref{fig:2_tau10E5_phi83-90}, \ref{fig:2_tau10E6_phi83-90} and \ref{fig:2_tau10E7_phi83-90}.\\

@@

In the next subsection, we show only the results of the angle at which an observer sees the galaxy's angular momentum vector perpendicular to the line of sight. The purpose of this is only to decrease the number of plots in the document, as there is an analogous behavior for all viewing angles. \\

>>>>

We model spherical galaxy to facilitate the interpretation of the results, by deleting what would be another degree of freedom regarding a different geometry. Furthermore, this approximation is commonly used in the literature, as it explains a wide variety of observational features (\cite{Ahn03}, \cite{Verhamme06}, \cite{Dijkstra06}). \\

@@

In this model, the LAE has a bulk velocity corresponding to the superposition of rotation and outflows, as shown in Fig. \ref{fig:model}. The velocity components are written in Eqs. (\ref{eq:vx}) (\ref{eq:vy}) (\ref{eq:vz}). In these equations, $R$ is the radius of the sphere; $x$, $y$ and $z$ are the coordinates in a cartesian frame; and the $\mp$ signs in $v_x$ and $v_y$ indicate the direction of rotation, respectively. This rotation is a solid body rotation and its direction goes according to the right hand rule applied to the $\hat{k}$ unit vector. The outflow velocity is dependent on the position relative to the center of the galaxy, being it zero at the center and maximum at the edge of the sphere.\\

@@

If $x < 0$, the final frequency $\nu < \nu_{\rm \alpha}$. This translates to the final wavelength being larger than the \lya natural one, casing a redshift in frequency. If, on the contrary, $x > 0$, the photon suffers a blueshift in frequency. \\

@@

The initial emission of photons is taken at the center of the sphere for practicality due to the fact that both, center and off-center emissions, give analogous results. From here, $100000$ photons are emitted with the natural \lya wavelength starting to behave as described in sub-section \ref{subsec:photons_path_clara} (See Fig. \ref{fig:radiative_transfer}). When each photon is re-emitted, its new wavelength depends on the H atom's velocity (both thermal and bulk) and direction (both initial and final). However the photon's new direction of propagation is random in the rest frame of the atom. \\

@@

%The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation. Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires. Depending on \tauh, we need a different number of processors in order to minimize the time demands. For \tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively. The running times go from 7 hours to 1 week approximately, depending on the case. \\

@@

\subsection{Parameters' Initial Values}
In order to define the ranges of \tauh, \vrot and \vout, it is necessary to refer to observational constraints. The common values for typical LAEs are in Tab. \ref{tab:values}. We run CLARA's modified version for all the permutations of these 3 parameters. \\

@@

The resulting sets of spectra are in the Appendix \ref{ap:results_figures} in Figs. \ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$. We do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms. For this reason, we limit \vout to a maximum value of 25 \kms in all the plots. Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{ap:additional_figures} in Figs. \ref{fig:2_tau10E5_phi83-90}, \ref{fig:2_tau10E6_phi83-90} and \ref{fig:2_tau10E7_phi83-90}.\\

@@

In the next subsection, we show only the results of the angle at which an observer sees the galaxy's angular momentum vector perpendicular to the line of sight. The purpose of this is only to decrease the number of plots as there is an analogous behavior for all viewing angles. \\
